---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#install.packages("factoextra")
#install.packages("kableExtra")

```
Load Packages
```{r}
library(tidyverse)  # data manipulation
library(cluster)    # clustering algorithms
library(factoextra)
library(pacman)
p_load(tidyverse,vegan,caret)
library(kableExtra)
```
Load Dataset
```{r}
Happy <-read.csv("/Users/dominiquedesmarattes/Documents/Senior Year School Docs/BMGT431/Group_Project/World_Happiness_Report_2022.csv", header=T)
```
Remove all columns that are not the 6 factors we're focusing on
```{r}
Happy <- Happy[,-c(1,3:6) ]
```
Rename Columns
```{r}
Happy <- Happy%>%
  rename(
    GDP = 'Explained.by..GDP.per.capita',
    Social.Support = 'Explained.by..Social.support',
    Life.Expectancy = 'Explained.by..Healthy.life.expectancy',
    Free.Choices = 'Explained.by..Freedom.to.make.life.choices',
    Generosity = 'Explained.by..Generosity',
    Corruption = 'Explained.by..Perceptions.of.corruption'
  )
```
##Hierarchical Clustering
Step 1: Normalizing
```{r}
#convert to a data frame for step 2 (row names)
Train.norm <- as.data.frame(Happy)

# use preProcess() from the caret package to normalize the data
norm.values <- preProcess(Happy[, 2:7], method=c("center", "scale"))

#We then use the norm.values to scale both the training and the validation data
Train.norm[, 2:7] <- predict(norm.values, Train.norm[, 2:7])

summary(Train.norm)

```
Step 2: Make the Country names the row names
```{r}
# set row names to the university column
row.names(Train.norm) <- Train.norm[,1] 

#drop Universities from the data
Train.norm <- Train.norm[,-1]

```
We utilize the hclust() function for the clustering. the hclust method needs a distance matrix and a cluster function as argument.
```{r}
#create the distance matrix
d <- dist(Train.norm, method = "euclidean")

#run the clustering
hc2 <- hclust(d, method = "average")
plot(hc2, hang = -1, ann = FALSE)

groups <- cutree(hc2, k=4) # cut tree into 12 clusters
# draw dendogram with red borders around the 12 clusters
rect.hclust(hc2, k=4, border="red") 
```
Finally, lets get a list of membership
```{r}
cutree(hc2, k = 4)
```
Add in the membership and summarize clusters
```{r}
#combine cluster membership with the orignial data
Happy <-cbind(Happy,clusterNum=cutree(hc2,k=4))

#summarize the data
Happy %>% group_by(clusterNum) %>% summarize(
  AvgGDP=mean(GDP),
  AvgSS=mean(Social.Support),
  AvgLE=mean(Life.Expectancy),
  AvgF=mean(Free.Choices),
  AvgG=mean(Generosity),
  AvgC=mean(Corruption)
)
```
##Kmeans Clustering
Step 1: Normalizing
```{r}
#convert to a data frame for step 2 (row names)
Train.norm <- as.data.frame(Happy)

# use preProcess() from the caret package to normalize the data
norm.values <- preProcess(Happy[, 2:7], method=c("center", "scale"))

#We then use the norm.values to scale both the training and the validation data
Train.norm[, 2:7] <- predict(norm.values, Train.norm[, 2:7])

summary(Train.norm)

```
Step 2: Make the Country names the row names
```{r}
# set row names to the university column
row.names(Train.norm) <- Train.norm[,1] 

#drop Universities from the data
Train.norm <- Train.norm[,-1]
```
Step 3: Run the clustering
```{r}
km <- kmeans(Train.norm, 4, nstart = 25)
```
Step 4: Characterize the clusters
```{r}
km_cluster <- km$cluster
#number of records in each cluster
km$size
#total distance between records in each cluster
km$withinss
#table of cluster centers of each cluster for each of the 6 factors of happiness 
km$centers
```
Step 5: List Clusters as Dataframe
```{r}
## list of cluster assignments
km_cluster <- order(km$cluster)
clusters <- data.frame(km$cluster[km_cluster])

```
Step 6: Visualize the output 
```{r}
#for easier reproducibility
MyNumClusters=4;
MyNumVariables=6;

# plot an empty scatter plot
plot(c(0), xaxt = 'n', ylab = "", type = "l", 
     ylim = c(min(km$centers), max(km$centers)), xlim = c(0, MyNumVariables))

# label x-axes
axis(1, at = c(1:MyNumVariables), labels = names(Train.norm))

# plot centroids (note if you are using more than eight cluster you need to extend the if statement)
for (i in c(1:MyNumClusters))
lines(km$centers[i,], lty = i, lwd = 2, col = ifelse(i %in% c(1, 3, 5, 7),
                            "black", "dark grey"))

# name clusters
text(x = 0.5, y = km$centers[, 1], labels = paste("Cluster", c(1:MyNumClusters)))
```
Step 5: Determine right k
```{r}
#Note that tot.withinss gives us the total within-cluster sum of squares, i.e. sum(withinss).

# Compute and plot wss for k = 1 to k = 15
k.values <- 1:15
wss_values<-c()

# extract wss for 2-15 clusters
for (i in 1:15){ 
  wss_values[i] <-kmeans(Train.norm, i, nstart = 10 )$tot.withinss
}

plot(k.values, wss_values,
       type="b", pch = 19, frame = FALSE, 
       xlab="Number of clusters K",
       ylab="Total within-clusters sum of squares")
```

```{r}
fviz_nbclust(Train.norm, kmeans, method = "wss", k.max = 15)
```

```{r}
fviz_cluster(km, data = Train.norm)
```

```{r}
Train.norm %>%
  mutate(Cluster = km$cluster) %>%
  group_by(Cluster) %>%
  summarise_all("mean")
```
Add in the membership and summarize clusters
```{r}
#combine cluster membership with the orignial data
Happy <-cbind(Happy,km$cluster)
```
Rename km$clusters column
```{r}
Happy <- Happy%>%
  rename(
    Cluster = 'km$cluster')
```
Create Pivot Table in Excel
```{r}
#to analyze in Excel, we export the information 
write.csv(Happy,"Happy.csv", row.names = FALSE)
```
Create Pivot Table in R
```{r}
#summarize the data
Happy_Table <- Happy %>% group_by(Cluster) %>% summarize(
  AvgGDP=mean(GDP),
  AvgSS=mean(Social.Support),
  AvgLE=mean(Life.Expectancy),
  AvgF=mean(Free.Choices),
  AvgG=mean(Generosity),
  AvgC=mean(Corruption)
)
```

```{r}
write.csv(Happy_Table,"Happy_Pivot_Table.csv", row.names = FALSE)
```

```{r}
Happy2 <-read.csv("/Users/dominiquedesmarattes/Documents/Senior Year School Docs/BMGT431/Group_Project/World_Happiness_Report_2022.csv", header=T)
```

```{r}
Rank <- Happy %>% 
  cbind(Happy,Happy2$RANK)
```

```{r}
Rank <- Rank[,c(16,17) ]
```


```{r}
Rank <- Rank%>%
  rename(
    Rank = 'Happy2$RANK')
```

